### 2023-03-16 outputs/paper/paper.qmd
update paper

---
title: "GSS Education Analysis (Working Title)"
subtitle: "GSS Education Analysis (Working Subtitle)"
author: 
  - Emily Kim
  - Huda Sahaf
  - Chloe Thierstein
thanks: "Code and data are available at: https://github.com/cthierst/gss_education_analysis.git" 
date: "March 16, 2023"
date-format: "D MMMM YYYY"
abstract: "abstract (work in progress)"
format: pdf
toc: TRUE
number-sections: TRUE
bibliography: references.bib
---

```{r}
#| include: false

#### Load Library ####
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(modelsummary)
library(kableExtra)
library(knitr)
library(dplyr)

#### Import cleaned_gss_2021.csv ####
cleaned_gss_2021 <-
  read.csv(here::here("./inputs/data/2021_data/cleaned_gss_2021.csv"))

```

# Introduction {#sec-introduction}

# Data and methods {#sec-data}

## Data Management {#sec-first-point}

This paper utilzes the R statistical programming language [@citeR], along with several packages, tidyverse [@citetidyverse], here [@citehere], janitor [@citejanitor],and dplyr [@citedplyr]. All figures in this paper have been created using the packages ggplot2 [@citeggplot2] and the tables have been created using knitr [@citeknitr] and kableExtra [@citekablextra]. The color styles in graphs have been created using the RColorBrewer packages [@citercolorbrewer].

## Source {#sec-second-point}

All data within this paper are extracted from the 2021 United States General Social Survey (GSS). This survey consists of a series of nationally representative cross-sectional interviews and collects data on contemporary American society in order to explain trends in opinions, attitudes and behaviors and monitor these trends **\[citegsscodebook\].** Since it began tracking trends in public opinion in 1972, the GSS has used in-person data collection as its primary method of data collection but in 2021, it moved to an address-based sampling method with a focus on web and web-based self-administered questionnaires **\[citegsscodebook\]**. The 2021 GSS was conducted from December 1, 2020 to May 3, 2021, and was processed in accordance with standard NORC procedures **\[citegsscodebook\].**

## Sampling {#sec-third-point}

The 2021 GSS samples adults over the age of 18 in the United States who do not live in institutional housing **\[citegsscodebook\].** This GSS was released in three batches, the first consisted of 10,091 addresses, the second consisted of 10,000 addresses and the third consisted of 7,500 addresses **\[citegsscodebook\]**. From these three releases, 1,271 completed the survey from the first batch, 1,391 completed the survey from the second batch, and 1,069 completed the survey from the third batch **\[citegsscodebook\].** Meaning that there was an overall response yield of 13.5% or 3,731 complete responses from the 27,591 total surveys released **\[citegsscodebook\].**

## Key Features {#sec-fourth-point}

This paper is assessing our estimand of 1,567 responses of the 3,731 complete responses received which represents 42% of the complete respondents and only 5.7% of the total surveys released. Respondent's to the 2021 GSS are males or female US citizens, not living in institutional housing, who are over the age of 18 **\[citegsscodebook\]**. In our analysis we removed 2,164 responses to account for unanswered or not applicable responses. We did this to ensure that the data used in this paper is complete and representative of all variables. The variables we selected for our analysis can be seen in @tbl-1 and their measurement levels can be seen in @tbl-2. Responses were measured using likert scales which measured respondents opinions to the questions being asked.

```{r}
#| include: false

#### Create Variable Description Tibble ####

 variable_table <-
  tibble(
    "Variable" = c(
      "class",
      "degree",
      "finrela",
      "satjob",
      "satfin"
    ),
    "Variable Description" = c(
      "Self-ascribed class of the respondent",
      "Highest level of education achieved by respondent",
      "Self-ascribed ranking of respondnet when compared to other US households",
      "Level of satisfaction with the work respondents do",
      "Level of satisfaction with the job respondents have"
    )
  )
```

```{r}
#| include: false

#### Create Variable Description Tibble ####

 variable_measure_table <-
  tibble(
    "Variable" = c(
      "class",
      "degree",
      "finrela",
      "satjob",
      "satfin"
    ),
    "Variable Measurement" = c(
      "Lower Class, Working Class, Middle Class, Upper Class",
      "Less Than High School, High School, Associate / Junior College, Bachelors, Graduate",
      "Far Below Average, Below Average, Average, Above Average, Far Above Average",
      "Very Satisfied, Moderately Satisfied, A Little Dissatisfied, Very Dissatisfied",
      "Pretty Well Satisfied, More or Less Satisfied, Not Satisfied at All"
    )
  )
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-1
#| tbl-cap: "Variable Descriptions"

#### Add Variable Table ####
variable_table |>
  kable(booktabs = TRUE) |>
  kable_styling(font_size = 9, latex_options = "hold_position")

```

```{r}
#| echo: false
#| warning: false
#| label: tbl-2
#| tbl-cap: "Variable Measurements"

#### Add Variable Table ####
variable_measure_table |>
  kable(booktabs = TRUE) |>
  kable_styling(font_size = 9, latex_options = "hold_position")
```

## Bias and ethics {#sec-fifth-point}

# Results {#sec-results}

```{r}
#| include: false

#### Ordering Highest-Level of Education to be Easier to Read in Graphs####

cleaned_gss_2021$degree <- 
  factor(cleaned_gss_2021$degree, 
         levels = c("Less Than High School", "High School", "Associate / Junior College", "Bachelors Degree", "Graduate Degree"))
```

```{r}
#| include: false

#### Ordering Job Satisfaction to be Easier to Read in Graphs####

cleaned_gss_2021$satjob <- 
  factor(cleaned_gss_2021$satjob, 
         levels = c("Very Satisfied", "Moderately Satisfied", "A Little Dissatisfied", "Very Dissatisfied"))
```

```{r}
#| include: false

#### Ordering Self-Ascribed Financial Wellness Compared to Other US Families ####

cleaned_gss_2021$finrela <- 
  factor(cleaned_gss_2021$finrela, 
         levels = c("Far Above Average", "Above Average", "Average", "Below Average", "Far Below Average"))
```

```{r}
#| include: false

#### Ordering Self-Ascribed Social Class of Respondents####

cleaned_gss_2021$class <- 
  factor(cleaned_gss_2021$class, 
         levels = c("Lower Class", "Working Class", "Middle Class", "Upper Class"))
```

```{r}
#| include: false

#### Ordering Financial Satisfaction####

cleaned_gss_2021$satfin <- 
  factor(cleaned_gss_2021$satfin, 
         levels = c("Not Satisfied at All","More or Less Satisfied", "Pretty Well Satisfied"))
```

When looking at @fig-1, we can see that the majority of respondents have only completed their high school diploma with the next highest response category being bachelor's degree and the third being graduate degrees. Generally respondents who identified as having their bachelor's degree or graduate degree reported being more likely to be "above average" and "far above average" than respondents who reported having an associate or junior college degree or lower. Interestingly, a large portion of respondents with higher levels of education still reported themselves as "average" when comparing themselves to other US families. Yet in @fig-1, we can see that at almost all levels of education, the majority of respondents rank themselves as "average" or lesser. The only group in which this is false, is in the group of respondents who have graduate degrees. In this group the majority rank themselves as "above average" or higher. Additionally, the rate at which those with less than a high school diploma and high school diplomas self-report ranking when compared to other US households decreases substantially at every measuring point as ranking increases. Importantly, we can see that at each level of education, there are respondents who rank themselves as "above average" making it appear that there may be other factors that are relevant in how respondents rank themselves.

```{r}
#| echo: false
#| warning: false
#| label: fig-1
#| fig-cap: "Respondent's Self-Ascribed Financial Ranking Against Other US Families by Their Highest Level of Education Achieved"

#### Creating Graph to Compare Respondent's Self-Ascribed FinancialRanking Against Other US Families by Their Highest Level of Education Achieved ####

cleaned_gss_2021 |>
  ggplot(mapping = aes(y = degree, fill = finrela)) +
  geom_bar() + 
  theme_classic() +
  labs(y = "Highest Level of Education Achieved",
       x = "Number of Respondents",
       fill = "Rating Against Other US Families")+
  scale_fill_brewer(palette = "Pastel1")+
  theme(legend.title = element_text(size = 6),
        legend.text= element_text(size=6))
```

@fig-2 demonstrates that all respondents are likely to rank themselves as "moderately satisfied" or higher in job satisfaction regardless of level of education. Interestingly we can see that there is a level of dissatisfaction at from each. group of respondents alluding to the idea that education level may not guarantee job satisfaction. This is especially clear when comparing those with bachelor's degrees to those with high school diplomas as each have similar levels of satisfaction and dissatisfaction.

```{r}
#| echo: false
#| warning: false
#| label: fig-2
#| fig-cap: "Respondent's Degree of Job Satisfaction by Their Highest Level of Education Achieved"

#### Creating Graph to Respondent's Degree of Job Satisfaction by Their Highest Level of Education Achieved ####

cleaned_gss_2021 |>
  ggplot(mapping = aes(y = degree, fill = satjob)) +
  geom_bar() + 
  theme_classic() +
  labs(y = "Highest Level of Education Achieved",
       x = "Number of Respondents",
       fill = "Job Satisfaction")+
  scale_fill_brewer(palette = "Pastel1")+
  theme(legend.title = element_text(size = 6),
        legend.text= element_text(size=6))
```

@fig-3 demonstrates that most respondents are "more or less satisfied" or higher when discussing their degree of financial satisfaction. Importantly, we can see that those with higher levels of education have higher rates within their group of being "pretty well satisfied" with their financial standing, than those with lower levels of education. Additionally, we can see that those with associate or junior college level education are the most balanced in all levels of financial satisfaction, alluding to the idea that a specific type of degree is influential in how respondents may respond.

```{r}
#| echo: false
#| warning: false
#| label: fig-3
#| fig-cap: "Respondent's Report of Financial Satisfaction by Their Highest Level of Education Achieved"

#### Creating Graph to Compare Respondent's Report of Financial Satisfaction by Their Highest Level of Education Achieved ####

cleaned_gss_2021 |>
  ggplot(mapping = aes(y = degree, fill = satfin)) +
  geom_bar() + 
  theme_classic() +
  labs(y = "Highest Level of Education Achieved",
       x = "Number of Respondents",
       fill = "Financial Satisfaction")+
  scale_fill_brewer(palette = "Pastel1") +
  theme(legend.title = element_text(size = 8),
        legend.text= element_text(size=8))
```

By looking at @fig-4 we can see that those with graduate and bachelor's degrees are much more likely to describe themselves as "upper class" than "lower class." Interestingly, in all levels of education, most respondents ranked themselves as either "working class" or "middle class" this could be influenced by American idealism of "hard-work" or how respondent's see themselves among their peers which is also subjective. Additionally, only a very small number of those with associate or junior college degrees and lower education levels describe themselves as "upper class," ranging from 0 to 10% of respondents from these groups.

```{r}
#| echo: false
#| warning: false
#| label: fig-4
#| fig-cap: "Respondent's Self-Ascribed Social Class by Their Highest Level of Education Achieved"

#### Creating Graph to Compare Respondent's Self-Ascribed Financial Ranking Against Other US Families by Their Self-Ascribed Social Class ####

cleaned_gss_2021 |>
  ggplot(mapping = aes(y = degree, fill = class)) +
  geom_bar() +
  theme_classic() +
  labs(
    y = "Highest Level of Education Achieved",
    x = "Number of Respondents",
    fill = "Social Class") +
  scale_fill_brewer(palette = "Pastel1") +
  theme(legend.title = element_text(size = 8),
        legend.text= element_text(size=8))
```

# Discussion {#sec-discussion}

## The American Dream {#sec-the-american-dream}

The United States is built upon individualism, which trickles into their definitions of financial security, happiness, and overall fulfillment and satisfaction in life. Well-being of individuals within the United States therefore, is tracked and monitored through a variety of different factors. The aim of this investigation is to determine whether higher educational achievement influences respondent\'s perceptions of their social class, degree of financial satisfaction and job satisfaction. These factors were chosen because of their significance with regards to the American Dream, a concept that is built off a linear pattern of behavior in terms of economic security. Hustle culture is popular within the United States as an extension of the American Dream, because the idea that an individual must work hard in order to reap the benefits in terms is what drives culture within America.

According to Economic Security and the American Dream, the rewards of the American Dream included owning a home, having access to quality health care, having a job that enables an individual to support their family, and having a secure and stately retirement [@citerankhirschl]**.**These are the fundamentals in terms of how Americans generally define their well-being in society. However, no matter how linear an American\'s personal development or educational journey may be, it is simply impossible to be able to guarantee the rewards that the American Dream has to offer. But the ideals of living in the United States, has a powerful impact on culture, and how individuals view themselves and their position within society, but how they fare in comparison to those around them.

The assumption in general is a higher degree of education translates to an overall greater quality of life or well-being, greater job satisfaction, higher social class, etc. However, we found that this was not necessarily the case. Overall, out of our participants that completed the survey, the largest group had only completed high school, and the second largest group had completed a bachelor\'s degree. The level of education was compared to a few variables: self-ascribed social class, financial satisfaction,  job satisfaction, and financial ranking against other families in the US.

## Social Class {#sec-social-class}

However, in some cases, the positive correlation between higher education and overall subjective well-being is more prominent - resulting in significant tension in our data of consideration. Researchers have investigated the relationship between education and one\'s subjective well-being, with several studies indicating that a higher level of education may be positively associated with greater levels of happiness. This relationship may be explained by the fact that education can impact one\'s self-ascribed social class, which in turn can influence their notion of happiness. Higher education is assumed to provide individuals with better job opportunities, higher income, better access to resources, and greater social mobility - placing people on the path towards a successful American Dream and resulting in a higher sense of one\'s social status. As a result, these individuals might experience greater life satisfaction, which contributes to their overall sense of happiness and security.

In a 2020 article published in the Journal of Personality and Social Psychology, authors Siyu Yu and Steven L. Blader examined the impact of status and social class on subjective well-being (SWB). They found that often times social class addresses fundamental psychological needs, such as autonomy, mastery, respect, and occupational prestige, and they proposed that individuals in higher social classes experience greater levels of status and power, leading to higher SWB **[@citeyublader].** Conversely, individuals in lower social classes feel lower levels of status and power, resulting in lower levels of SWB. This insight is evident in @fig-3, where we found that individuals holding Bachelors and Graduate Degrees are proportionally more satisfied with their financial situation, with the large majority selecting \"Pretty Well Satisfied\" and \"More or Less Satisfied\" than those selecting \"Not Satisfied at All\". Out of the 426 responses indicating a \"Pretty Well\" financial satisfaction, 69% of responses were individuals holding Bachelor and Graduate degrees while only 31% held Associate/Junior college, highschool, and less than highschool levels of education. In comparison, those who only completed up until the high school level of education have the most responses of \"Not Satisfied at All\". 70% out of the 390 \"Not Satisfied At All\" responses were individuals who completed highschool education or lower, while only 30% of those holding Bachelor and Graduate degrees indicated a low financial satisfaction.

In addition, when looking at @fig-4, we found that there are more individuals holding Bachelor and Graduate degrees who have identified themselves as Upper and Middle class in proportion to those with lower levels of education. 89% of those who selected \"Upper Class\" and 68% of \"Middle Class\" held Bachelor and Graduate degrees. Among the 676 responses for \"Working Class\", 471 of those held high school level of education and lower, making up 70% of the responses. In sum, we found that both the research and data point to a link between higher social class and greater perceived levels of life satisfaction.

Overall, there seems to be a significant relationship between the data in education, social class, and one\'s perceived well-being. Education may serve as an important factor in determining an individual\'s social class as it can unlock opportunities for resources and mobility, which in turn can positively impact their sense of status and overall happiness. At the same time, while higher levels of education may be generally associated with greater levels of overall well-being, this relationship may not always be a cause and effect correlation. Subjective well-being may be further influenced by other factors such as personal values, cultural background, and social support. It is important to consider these important nuances when examining the data and making assumptions in order to develop a more comprehensive understanding of how variables interact and impact individuals\' lives. This will be further explored in the weaknesses, ethics, and bias section of this paper.

## Financial Satisfaction {#sec-financial-satisfaction}

Financial satisfaction as a result of hard work is the core principle within the American Dream, and as a result, Americans are known for their work culture. Researchers have tried to investigate the differences between work culture within Europe as well as within the United States and have narrowed it down to a variety of factors that contribute to the culture as it is today. Despite the fact that the economies and societal frameworks of Europe and America are very similar, the approach to work within the United States is very different. The number of hours worked per year for the average American was about 17% higher than the hours per year for the average European. [@citeeuropevacay] For this reason, financial satisfaction and social class are very prevalent within the American definition of well-being. Job satisfaction ties very closely this idea in the pursuit of the American Dream.

The second largest group of participants who had completed the survey had completed a bachelor's degree. In general, while a higher level of education improves access to better job opportunities, the process is not as linear as it may seem. According to @fig-1, while the number of individuals who have a graduate degree is not as high as the group that has completed a bachelor\'s degree, the reported financial ranking against other families was proportionately very similar for both groups. The majority for both groups claimed to be average or above average. There was still a significant number of individuals who had chosen below average, even for the group with the highest level of education.

A similar notion exists when comparing an individuals level of education with their job satisfaction. According to @fig-2,as the education level rises so does the number of people who are proportionately more satisfied with their jobs. However, job satisfaction is fleeting, it is simply a reflection of the current state, and not an overall indication of cumulative job satisfaction. According to Figure 3, financial satisfaction rises with an increased level of education, but the group that claims dissatisfaction is a significant number in the bachelor\'s and graduate degree groups. It remains a minority, but it is significant nonetheless. This points towards the un-attainability of the American Dream, in that despite the greater level education and increased access that comes with it, it does not guarantee increased job satisfaction, or a higher financial satisfaction.

## Limitations and weaknesses {#sec-limitations-and-weaknesses}

There are several limitations to this paper. One weakness is that our insights rely on data from the General Social Survey (GSS), which is a cross-sectional survey that collects data at a single point in time. Although it can be helpful in investigating associations between variables and developing hypotheses, the GSS is not capable of establishing causation between variables and has potential for selection bias if participants are not representative of the population of interest [@citestrengthsweaknesses]. In addition, due to changes in the questions asked over time, the GSS has limited capacity to make exact comparisons - as in the data analyzed in our study, which was collected during a period when the survey was conducted entirely online, which necessitated modifications to the question format in order to accommodate the new platform. Above all, due to the nature of the data being self-reported responses, it may not fully capture the dynamic and complicated nature of the ways in which various levels of education affect one\'s overall well-being. The General Social Survey responses are subject to influence by a myriad of external factors including recall bias, in which participants may have left out important details when reporting about their subjective experiences.

Another potential ethical concern is that our paper may inadvertently reinforce the concept of the American Dream, which suggests that higher education is a key factor to achieving happiness and success in life. This concern is particularly relevant in @sec-social-class, where we analyze the data as-is without taking into account the presence of systemic inequalities such as racism and economic inequality that can significantly limit access to one\'s education and social mobility. Moreover, the sample survey in the GSS may not fully represent the experiences of more marginalized groups, which point to the need for further research and survey strategies that incorporate more diverse perspectives.

In sum, while our paper provides valuable insights on the relationship between education and one\'s perceived well-being, it is important to consider the limitations, ethical concerns, and potential biases in the data and interpretation of our insights and findings.

# Future Research {#sec-fur-reseach}

More research into this topic is important as it helps us gain a more substantial understanding of American ideals and how certain indicators may influence the rate at which individuals are achieving them, or not achieving them. Further research into this topic should explore more specific questions about attitudes towards work, type of work, income, and comparing respondent situations to close acquaintances to better understand American perspectives. We have provided a supplementary survey in @sec-appendix that would further build on the US GSS data in relation to our topic. Additionally, future research should be conducted over a period of time using the same survey and methodology to collect obtain information that is verifiable and encompasses a larger population of the United States.

\newpage

\appendix

# Appendix {#sec-appendix}

## Supplementary Survey

The supplementary survey is available: [here](https://forms.gle/vUCXy927cDnpx5UL9)

```{r, out.width = "500px"}
#| echo = FALSE

#### Import .png Files of Supplementary Survey ####

knitr::include_graphics("./supplementary_survey/suppl_surv_01.png")
knitr::include_graphics("./supplementary_survey/suppl_surv_2.png")
knitr::include_graphics("./supplementary_survey/suppl_surv_3.png")
knitr::include_graphics("./supplementary_survey/suppl_surv_4.png")
knitr::include_graphics("./supplementary_survey/suppl_surv_5.png")
knitr::include_graphics("./supplementary_survey/suppl_surv_6.png")

```

\newpage

# References